# FLightR analysis example of Black tailed godwit geolocator

## Install package
Note that we used FLightR 0.4.5 version, and running this example in later versions may provide somehow different results.
To install the latest CRAN version try
```{r, eval = F}
install.packages('FLightR')
```
To run version 0.4.5 which we use in this example try 
```{r, eval = F}
library(devtools)
install_github('eldarrak/FLightR@0.4.5')
```
The latest version is available here:
```{r, eval = F}
install_github('eldarrak/FLightR')
```
Load package
```{r}
library(FLightR)
```
## 1. Data preparation and import
The analysis of geolocation data begins with defining the twilight events, i. e. sunrises and sunsets. This procedure is not implemented in FLightR. Please, use appropriate functions in other R packages, e.g. BAStag, twGeos or GeoLight [Lisovski et al. 2012a](#lisovski_2012a), or online service http://tags.animalmigration.org for this purpose. In this software, user guides semi-automatically search for the twilight times and then visually explores the light data, manually removing twilights defined apparently wrong or around which the light pattern is atypical or systematically biased, e.g. when a bird enters or exits a cavity during twilight ([Rakhimberdiev et al. 2016](#rakhimberidev_jab_2016)). An example of the routine for the twilight detection in BAStag is available as [appendix A4](https://github.com/eldarrak/FLightR/blob/master/examples/Black-Tailed_Godwit_JAB_example/A4_BAStag_routine.Rmd) to [Rakhimberdiev et al. 2016](#rakhimberdiev_jab_2016).

FLightR works with the ‘TAGS’ format, which is an agreed general annotation of twilight data by National Centre for Ecological Analysis and Synthesis working group ‘Establishing an open-source animal-tracking analysis platform for archival geolocators https://www.nceas.ucsb.edu/featured/bridge. A ‘TAGS’ file is a 'CSV' file containing the following fields: 

* `datetime` – date and time in ISO 8601 format e.g. 2013-06-16T00:00:11.000Z;
* `light` – light value measured by tag;
* `twilight` – assigned by the software numeric indication of whether the record belongs to sunrise (1), sunset (2) or none of those (0);
* `excluded` – indication of whether a twilight was excluded during manual inspection (logical, `TRUE | FALSE`);
* `interp` - indication of whether the light value at twilight was interpolated (logical, `TRUE | FALSE`). 
The fields `excluded` and `interp` may have values of `TRUE` only for `twilight > 0`. 
The online service http://tags.animalmigration.org saves data in the TAGS format. In the R packages GeoLight and BAStag or twGeos, the annotated twilight data need to be exported to TAGS, for which the functions in the FLightR (`GeoLight2TAGS`,  `BAStag2TAGS` or `twGeos2TAGS`) can be used.

We will use the TAGS formatted data provided as Appendix A3 to [Rakhimberdiev et al. 2016](#rakhimberdiev_jab_2016) also available on the authors’ GitHub page:
```{r, eval=FALSE, tidy=FALSE}
download.file('https://git.io/vrJgv', 'example_TAGS_format.csv')
```
The function `get.tags.data` reads comma separated file in TAGS format, detects tag type, checks whether the light data are log-transformed, transforms them back from the log scale if needed an object, containing 

1. the recorded light data,
2. the detected twilight events,
3. light level data at the moment of each determined sunrise and sunset and around them (24 fixes before and 24 after it events into an object of two lists
4. technical parameters of the tag, added automatically, unless preset by the user.

Currently finction recognizes data from all common tag types: mk tags (produced by British Antarctic Survey, Lotek, and Migrate Technology Ltd.), Intigeo tags (Migrate technology Ltd.) and GDL tags (Swiss Ornithological Institute).
```{r, eval=FALSE, tidy=FALSE}
Proc.data<-get.tags.data("example_TAGS_format.csv")
```
 
## 2. Calibration

All geolocators measure light levels with different precision, and calibration is needed to find the relationship between the measured and real light levels for each device. This relationship is depicted by the calibration parameters (slopes), which are calculated based on the data recorded in known (calibration) locations.  Such locations can be the ones, where the animal was tagged, recaptured or observed. FLightR works with so-called template fit observational model [Ekstrom 2004, 2007](#ekstrom_2004). This model compares log of light irradiance measured at location with log of irradiance expected there according to astronomical position of the sun [Rakhimberdiev et al. 2015](#rakhimberdiev_movecol_2015).

The calibration parameters are based on the data recorded during calibration periods. To define these periods, user needs to create a data frame in which geographic coordinates of the calibration location (usually, the location, where the tag was deployed on an animal), and start and end date of the calibration period are specified in the following columns:
* `calibration.start` (POSIXct format)
* `calibration.stop` (POSIXct format)
* `lon` (numeric)
* `lat` (numeric)
The data frame contains as many rows as many distinct calibration periods are in the track. 

```{r, eval = F}
Calibration.periods<-data.frame(
        calibration.start=as.POSIXct(c(NA, "2014-05-05")),
        calibration.stop=as.POSIXct(c("2013-08-20", NA)),
        lon=5.43, lat=52.93) 
		#use c() also for the geographic coordinates, if you have more than one calibration location
		# (e. g.,  lon=c(5.43, 6.00), lat=c(52.93,52.94))
print(Calibration.periods)
```
In this example, we have two calibration periods (in the same location), in the beginning and at the end of the track. This is a common case, as the birds are oftentimes recaptured at the same location, where they were tagged. 

When multiple calibration locations are available, each of them has to be processed with the function `plot_slopes_by_location`.  In this case, in the ‘Calibration periods’ data frame, each row should refer to one calibration period. Compiling the data frame with multiple calibration locations, use `c()` also for the geographic coordinates (e. g., `lon=c(5.43, 6.00), lat=c(52.93,52.94)`).

Using the created data frame with the calibration periods and locations and the object with light data and twilight events (`Proc.data`), we estimate calibration parameters create a calibration object. This object will be used for the calculation of geographic positions later on.
```{r, eval = F}
Calibration<-make.calibration(Proc.data, Calibration.periods)
```
When there are more than one calibration periods, the parameter `model.ageing` can be set `TRUE` to account for tag ageing. In this case, the calibration parameters are calculated, based on the assumption that the calibration slope changes linearly (with time) between the calibration points. The obtained calibration object is further used for calculation of geographical positions.

### Find calibration periods for known calibration location
For the data recorded by a deployed on an animal tag, the exact period of its stay in a known location is usually not known, but it can be suggested from the data. For this, we obtain and plot calibration slopes for sunsets and sunrises for every day of the tracking period, based on the assumption that the tag remained in the same calibration location all the time. 

```{r, eval = F}
plot_slopes_by_location(Proc.data=Proc.data, location=c(5.43, 52.93))
```

Looking at the plot, we can define the time periods, during which the tag really was in the calibration location (recall, that we assume that the tag remained in this location all the time). Because calibration slopes reflect the adequacy of the light level measured by the device, they vary little, in time and between sunsets and sunrises, as long as the tagged animal stays in the calibration location, but become clearly diverse, when it moves away from it. Both patterns are clearly distinguishable at the plot. 

Play with `abline()` to find the proper boundaries for the calibration.
```{r, eval = F}
abline(v=as.POSIXct("2013-08-20")) # end of first calibration period
abline(v=as.POSIXct("2014-05-05")) # start of the second calibration period
```
### Find calibration location for known calibration period
When tagging happened in the High Arctic under polar day conditions or when bird moved away from the tagging site immedialtey after tagging and rood top calibration data are absent it is still possibe to obtain calibration parameters for a resident period at unknown location. FLightR approach to this problem is similar to Hill-Ekstrom calibration [
[Lisovski et al. 2012b](#lisovski_2012b) implemented in GeoLight [Lisovksi et al. 2012a](#lisovski_2012a). If bird is assumed to be resident at some period one can try:
```{r, eval=FALSE}
# ~ 15 min run time
Location<-find.stationary.location(Proc.data, '2013-07-20', '2013-08-20', initial.coords=c(10, 50))
```
Function will return coordinates of location for which spread of errors in slopes is minimal. Initial coordinates should be within a few thousand kilometers from the real location.

## 3. Spatial extent

Set up a spatial grid (50 X 50 km on default) `left`, `right`. `bottom` and `top` are specific boundaries for your bounding box. `distance.from.land.allowed.to.use` should be vector with length of two, first number is negative distance allowed to use while over land, and second is distance from land allowed to use while over water.
`distance.from.land.allowed.to.stay` should be vector with length of two, first number is negative distance where bird is allowed to be stationary, and second is distance from land allowed to fly over during twilight while over water.
```{r, eval = F}
Grid<-make.grid(left=-14, bottom=30, right=13, top=57,
   distance.from.land.allowed.to.use=c(-Inf, Inf),
   distance.from.land.allowed.to.stay=c(-Inf, Inf))
```
The resulting grid is a matrix with the columns: `lon` (longitude), `lat` (latitude) and `Stay` (probability of stay). The grid cells, which the animal presumably cannot use (e.g. locations over land for a marine animal), are excluded from the data, while the locations at which an animal cannot be stationary are given a low probability of stay. Using masks can side track model estimation to the local minima, so we recommend to initially run model without a mask, enable them for the second run and visually compare the results, to see if the model converges to a similar track.

## 4. Create proposal

Here we create a complex object, which will be used in the main run. It incorporates all the objects, created at earlier steps: the light data with the detected twilight events (Proc.data), the spatial parameters (Grid), geographic coordinates of the initial location, where the tracking has started (start), and the calibration parameters (Calibration).

```{r, eval = F}
# ~ 15 min run time
all.in<-make.prerun.object(Proc.data, Grid, start=c(5.43, 52.93), Calibration=Calibration)
```
## 5. Run particle filter

At this stage, results are calculated (coordinates, behavior, probability of migration).
Within the `run.particle.filter`, the following parameters can be defined: 
* `nParticles` - number of particles (1e4 is recommended for test and 1e6 for the analysis);
* `threads` - amount of parallel threads to use for the run default is -1 that means all available except one;
* `known.last` - TRUE if you know that in the end of data collection tag was in a known place (FALSE is the default option);
* `check.outliers` – FALSE by default. Set it TRUE if you wish on a fly outliers detection, we recommend use it if results have strong outliers.

```{r, eval = F}
nParticles=1e6
# ~ 45 min run time
Result<-run.particle.filter(all.in, threads=-1,
            nParticles=nParticles, known.last=TRUE,
		    precision.sd=25, check.outliers=F)
save(Result, file="Result.bltg.ageing.model.noOD.RData")
```
Now your results are saved in your work directory as an RData object.

## 6. Migration schedules

We can extract departure arrival estimates from the results we have got. 
First select grid points that are of interest. For example in the current data we are interested to figure out when our bird left the Netherlands. We will make  a boundary at 2&deg; longitude;
```{r, eval = F}
Index<-which(Result$Spatial$Grid[,1]>(2))
```
And now we estimate probabilities if being in the area for each twilight:
```{r, eval = F}
Arrivals.NL<-find.times.distribution(Result,Index)
Arrivals.NL
```
## 7. Result summary and plots

### Print summary on stationary and migration periods
```{r, eval = F}
stationary.migration.summary(Result, prob.cutoff = 0.1, min.stay = 3)
```
### Plot a simple map
```{r, eval = F}
map.FLightR.ggmap(Result)
```
### Plot lon lat graph
```{r, eval = F}
plot_lon_lat(Result)
```
### Plot utilization distribution
Plot space utilisation distribution for the wintering range 
```{r, eval = F}
plot_util_distr(Result, 
    dates=data.frame(as.POSIXct('2013-12-01'), as.POSIXct('2014-01-31')),
    add.scale.bar=TRUE, percentiles=0.5)
```

## References

1. Ekstrom, P. (2004). An advance in geolocation by light. Memoirs of the National Insitute of Polar Research, Special Issue, 58, 210–226.<a id="ekstrom_2004"></a>
1. Ekstrom, P. (2007). Error measures for template-fit geolocation based on light. Deep Sea Research Part II: Topical Studies in Oceanography, 54, 392–403.<a id="ekstrom_2007"></a>
1. Lisovski, S., Hahn, S. & Hodgson, D. (2012a). GeoLight - processing and analysing light-based geolocator data in R. Methods in Ecology and Evolution, 3, 1055–1059.<a id="lisovski_2012b"></a>
1. Lisovski, S., Hewson, C.M., Klaassen, R.H.G., Korner-Nievergelt, F., Kristensen, M.W. & Hahn, S. (2012b). Geolocation by light: accuracy and precision affected by environmental factors. Methods in Ecology and Evolution, 3, 603–612.<a id="lisovski_2012a"></a>
1. Rakhimberdiev, E., Winkler, D.W., Bridge, E., Seavy, N.E., Sheldon, D., Piersma, T. & Saveliev, A. (2015). A hidden Markov model for reconstructing animal paths from solar geolocation loggers using templates for light intensity. Movement Ecology, 3, 25.<a id="rakhimberdiev_movecol_2015"></a>
1. Rakhimberdiev, E., Senner, N. R., Verhoeven, M. A., Winkler, D. W., Bouten, W. and Piersma T. (2016) Comparing inferences of solar geolocation data against high-precision GPS data: annual movements of a double-tagged Black-Tailed Godwit. - Journal of Avian Biology 47: 589-596.<a id="rakhimberdiev_jab_2016"></a>
