<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{Bar-Tailed Godwit GeoLight analysis}
%\VignetteDepends{FLightR}
-->
# Intigeo tag on a Bar tailed godwit analysis example
# GeoLight analysis
Appendix A5 to 
Rakhimberdiev, E., Senner, N. R., Verhoeven, M. A., Winkler, D. W., Bouten, W. and Piersma T. 2016 Comparing inferences of solar geolocation data against high-precision GPS data: annual movements of a double-tagged Black-Tailed Godwit. - Journal of Avian Biology 000: 000-000.

We used [GeoLight](https://github.com/slisovski/GeoLight) 2.01, so you might want to install it if you want to get exactly the same results.
```{r, eval = F}
library(devtools)
install_github("SLisovski/GeoLight@Update_2.01")
library(GeoLight)
```
##Download data from GitHub
There are two files available in the directory: *.lux and *.csv. *.lux is the original file you get from Migrate Technology Ltd. This file does not have defined twilights. In the [appendix A4](https://github.com/eldarrak/FLightR/blob/master/examples/Black-Tailed_Godwit_JAB_example/A4_BAStag_routine.Rmd) we already defined twilights with [BAStag package] (https://github.com/SWotherspoon/BAStag) and saved them. You can define twilights by yourself or just download file with predefined format.
Let's now assume that you have done the previous step and have a [.csv](https://raw.githubusercontent.com/eldarrak/FLightR/master/examples/Black-Tailed_Godwit_JAB_example/A3_TAGS_format.csv) file. Now you can process these data:

```{r, eval = F, tidy=FALSE}
download.file(
 "https://raw.githubusercontent.com/eldarrak/FLightR/master/examples/Black-Tailed_Godwit_JAB_example/A3_TAGS_format.csv",
 "A3_TAGS_format.csv")
TAGS.twilights<-read.csv("A3_TAGS_format.csv", stringsAsFactors =F)
``` 

##Prepare data
```{r, eval=FALSE}
TAGS.twilights$light<-exp(TAGS.twilights$light) # this is needed because
                      # we log transformed data in convert.lux.to.tags()

TAGS.twilights$datetime<-as.POSIXct(TAGS.twilights$datetime,
                                    tz="UTC", format="%Y-%m-%dT%T")

# so we have to extract twilights
tw<-TAGS.twilights[TAGS.twilights$twilight>0 &  TAGS.twilights$excluded==0
 & TAGS.twilights$datetime<as.POSIXct("2014-05-18", tz="UTC"),]

# make pairs
tw$datetime[tw$twilight==2]<-tw$datetime[tw$twilight==2]-(300-60)

gl_twl<-data.frame(tFirst=tw$datetime[which(diff(tw$datetime)<20)],
                   tSecond=tw$datetime[which(diff(tw$datetime)<20)+1],
				   type=tw$twilight[which(diff(tw$datetime)<20)])

```
## Calibration
```{r, eval=FALSE}
#Calibrate the sun elevation angle with a known location
#designate first twilights for calibration
calib1 <- gl_twl[gl_twl$tFirst<=as.POSIXct("2013-08-05", tz="UTC"),] 
known.coord <- c(5.47, 53.04)

elev <- getElevation(calib1$tFirst,calib1$tSecond,
                     calib1$type,known.coord=known.coord,
					 plot=T) #get sun elevation angle 
```
##Generate the coordinates
```{r, eval=FALSE}

crds <- coord(twl=gl_twl, degElevation = elev)

#Add times to GeoLight crds
crds <- data.frame(Time = as.POSIXct(apply(cbind(gl_twl$tFirst,gl_twl$tSecond)
                                           , 1 , mean), tz="UTC", origin="1970-01-01"),
                   Lon  = crds[,1], Lat = crds[,2])

tripMap(crds[,c("Lon", "Lat")], xlim=c(-10, 10), ylim=c(0, 60))
```

## Estimate sites and migration schedules
```{r, eval=FALSE}
cL <- changeLight(gl_twl,  quantile = 0.95, days = 3) #0.95 does not work..

mS <- mergeSites(gl_twl, site = cL$site, degElevation = elev, distThreshold = 400)

siteMap(crds[,c("Lon", "Lat")], mS$site, type = "cross", hull = F, xlim = c(-12, 12), ylim = c(30, 58))

(schedule(tFirst = gl_twl$tFirst, gl_twl$tSecond, site = mS$site))
```
Result:
```
  Site             Arrival           Departure
1    a                <NA> 2013-08-20 11:55:25
2    b 2013-08-26 00:10:58 2014-04-20 00:26:08
3    c 2014-04-22 00:17:36 2014-04-28 11:59:32
4    d 2014-05-02 23:45:04                <NA>
```

```{r, eval=FALSE}
crds$Site <- mS$site

save(crds, file="GeoLight_coords.RData")
```
Now we will proceed to [FLightR workflow](https://github.com/eldarrak/FLightR/blob/master/examples/Black-Tailed_Godwit_JAB_example/A6_FLightR_analysis.Rmd).
