<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{lux - BAStags - FLightR workflow}
%\VignetteDepends{FLightR}
%\VignetteDepends{BAStag}
-->
# .lux - BAStags - FLightR workflow
## Install package
```{r}
require(devtools)

install_github("SWotherspoon/BAStag")
install_github("eldarrak/FLightR") # note the version
library(BAStag)
library(FLightR)
```

Here I use FLightR after initial twilight detection by BAStag.
First we start with downloading the .lux file [.lux](https://raw.githubusercontent.com/eldarrak/FLightR/master/examples/black-tailed%20godwit_Intigeo_tag_example/E391_driftadj.lux) file into your working directory. I assume you will name it 'E391_driftadj.lux'.
2. open R and process .lux file
```{r}
d.lux<-readMTlux("E391_driftadj.lux")
d.lux<- subset(d.lux,select=c("Date","Light"))
```
In the lux file light values go very high, so we should log transform data before selecting twilights.
```{r}
d.lux$Light<-log(d.lux$Light)
```
now we plot the image and select offset

```{r}
offset = 10
lightImage(d.lux, offset = offset, zlim = c(0, 12), dt = 300) # dt specifies the recording interval

threshold=2 # better use 2 for Intigeo tags if no strong reason for other value
		  
twl <- preprocessLight(d.lux, threshold, offset = offset,lmax=12) # only needs to be done once

```
Transfer BAStag output to TAGS format...
```{r}
library(FLightR)

TAGS.twilights<-BAStag2TAGS(d.lux, twl, threshold=2.5)

TAGS.twilights$light<-exp(lig.raw$light) # this is needed only if you log transformed in the beginning..

```
Now you can read these data by the FLightR data